---
title: "anthroDisturbance_DataPrep Manual"
subtitle: "v.`r SpaDES.core::moduleMetadata(module = 'anthroDisturbance_DataPrep', path = '..')$version`"
date: "Last updated: `r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 4
    html_preview: false
editor_options:
  chunk_output_type: console
---

# anthroDisturbance_DataPrep Module

```{r setup-anthroDisturbance_DataPrep, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold") ## change to eval = TRUE if all chunks are to be evaluated
```

#### Authors:

`r paste(as.character(SpaDES.core::moduleMetadata(module = "anthroDisturbance_DataPrep", path = "..")$authors), sep = ", ")`

## Module Overview

### Module summary

This is a data-preparation module that harmonizes different anthropogenic
disturbance datasets into the list-of-lists structure needed to generate
disturbance layers downstream. All the module needs is the metadata supplied in
the `disturbanceDT` object (a `data.table`); from it, the module downloads,
loads, reprojects, crops, and harmonizes each dataset, returning the
`disturbanceList` object.

The module was originally developed for the Northwest Territories (its built-in
default sample data is the union of BCR6 and NT1), but the structure is general.
This is the FOR-CAST fork used in the Yukon Northern Mountain Caribou project,
where the pipeline supplies Yukon inputs.

### Module inputs and parameters

The primary input is the `disturbanceDT` `data.table`, whose columns describe
where each dataset comes from and how to process it:

- `dataName`: groups the data by sector (e.g., Energy, Settlements, OilGas,
  Mining, Forestry, Roads);
- `URL`: download link for the specific dataset;
- `classToSearch`: exact polygon type/class to search for when picking from a
  dataset with multiple types. Even if a shapefile already holds only the needed
  data, specify a unique value here so each entry has a distinct name;
- `fieldToSearch`: the field in which `classToSearch` is found; if given, the
  spatial object is subset to `classToSearch`. Provide only if needed;
- `dataClass`: further details the data type (e.g., Settlements,
  potentialSettlements, otherPolygons, otherLines, windTurbines,
  potentialWindTurbines, hydroStations, oilFacilities, pipelines). A common class
  used to harmonize different datasets. Potential data classes (always prefixed
  `potential`) fall into three general types, later consumed by
  `anthroDisturbance_Generator`:
    1. Enlarging (e.g., potentialSettlements, potentialSeismicLines): the
       potential layer equals the current layer and is only buffered over time;
    2. Generating (e.g., potentialWind, potentialOilGas, potentialMineral,
       potentialCutblocks): the potential layer marks where structures may appear
       based on a rate;
    3. Connecting (e.g., potentialPipelines, potentialTransmission,
       potentialRoads): the potential layer needs the current transmission,
       pipeline, and road network; depends on what Generating produces;
- `fileName`: if the source is a `.zip` holding one or more shapefiles, which
  shapefile to use;
- `dataType`: layer format -- one of `shapefile` (`.shp`/`.gdb`), `raster`
  (`.tif`, converted to shapefile), or `mif` (read as a shapefile).

The other two inputs are the study area (`studyArea`) and a raster
(`rasterToMatch`) that defines the projection and spatial resolution.

The full list of module inputs:

```{r moduleInputs-anthroDisturbance_DataPrep, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_inputs <- SpaDES.core::moduleInputs("anthroDisturbance_DataPrep", "..")
knitr::kable(df_inputs)
```

Summary of user-visible parameters:

```{r moduleParams-anthroDisturbance_DataPrep, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_params <- SpaDES.core::moduleParams("anthroDisturbance_DataPrep", "..")
knitr::kable(df_params)
```

### Events

The module runs two events:

1. `init`: validates and coerces the inputs -- `studyArea` to a terra
   `SpatVector`, `rasterToMatch` to a `SpatRaster`, and `disturbanceDT` to a
   `data.table` -- reprojecting `studyArea` to match `rasterToMatch` when needed;
2. `loadAndHarmonizeDisturbanceDT`: builds the disturbance data. `createDisturbanceList()`
   loops over `dataName`, `dataClass`, and `classToSearch` to download each
   dataset, load it, reproject non-matching layers to the `rasterToMatch`
   projection, crop to the study area, and merge layers sharing a common
   `dataClass`. `harmonizeList()` then unifies sub-sectors that can be simulated
   together into the final `disturbanceList`.

### Plotting

This module does not schedule any plot events.

### Saving

This module does not schedule any save events. When `useSavedList = TRUE`, a
previously-saved `disturbances` list is reused instead of being rebuilt, which
saves time -- but confirm the saved objects are the expected ones.

### Module outputs

```{r moduleOutputs-anthroDisturbance_DataPrep, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_outputs <- SpaDES.core::moduleOutputs("anthroDisturbance_DataPrep", "..")
knitr::kable(df_outputs)
```

Both outputs are lists (general sector category) of lists (specific class /
sub-sector). In `disturbances` the inner `dataClass` may not be unique; in
`disturbanceList` each `dataName` x `dataClass` combination is unique after
harmonization, except for `potential` resources that need idiosyncratic
processing. Those are handled by a downstream module -- for the current
defaults, `potentialResourcesYT_DataPrep`. If no potential layer needs
modification or combination, that module can be skipped and `disturbanceList`
passed directly to `anthroDisturbance_Generator`.

### Links to other modules

Part of the anthropogenic-disturbance module collection, run in sequence with
`potentialResourcesYT_DataPrep` (idiosyncratic potential-resource processing)
and `anthroDisturbance_Generator` (forward generation of new disturbances). The
collection can be combined with landscape-simulation modules (e.g., `Biomass_core`)
and fire (e.g., `scfm`) and caribou modules to improve realism in forecasts.

### Getting help

- <https://github.com/FOR-CAST/anthroDisturbance_DataPrep/issues>
